---
title: "HW02"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
I will be using the gapminder data set for this exploration
```{r}
library(gapminder)
library(tidyverse)
```

First I will explore the data set to familiarize myself with it

```{r}
class(gapminder)
str(gapminder)
names(gapminder)
ncol(gapminder)
nrow(gapminder)
```
Now I can see that it is a tibble data.frame, I can see the column headers, and number of columns and rows.

Next I will explore a bit more information about countries and their populations

```{r}

summary(gapminder)


```
I have some basic information now about the values and distribution 

```


Next I will try some plots to visualize the data better

```{r}
ggplot(gapminder, aes(x=lifeExp, y=pop)) + 
  geom_point()
ggplot(gapminder, aes(gdpPercap, pop)) +
  geom_point()

ggplot(gapminder, aes(lifeExp)) +
  geom_histogram(bins=50)

ggplot(gapminder, aes(lifeExp)) +
  geom_histogram() +
  scale_x_log10()
```



Now I will compare one category to one numberical variable and see what looks best. 

```{r}
a <- ggplot(gapminder, aes(country, pop))

a + geom_boxplot()
a + geom_violin()
```

This is a lot of data to visualize this way so I will narrow it down to continent instead.
```{r}
ggplot(gapminder, aes(continent, pop)) +
  geom_boxplot()

```


Now I'm going to filter my data to look at countries with life expectancy greater than 50.

```{r}
gapminder %>% 
  filter(lifeExp > 50)
  
```
Now I am curious about just life exp and population in the different countries. 

```{r}

gapminder %>% 
select(country, lifeExp, pop)
```

Now I am just curious about countries in africa that have a life expentancy greater than 50.
```{r}
gapminder %>% 
  filter(lifeExp > 50) %>% 
  filter(continent == "Africa")
```


Now I will look at this data in terms of population and select that data frame to  do some plots. First I will assign my filter a variable and then try different plots to see what looks best.

```{r}
c <- gapminder %>% 
  filter(lifeExp > 50) %>% 
  filter(continent == "Africa")

c %>% 
  ggplot(aes(country,lifeExp)) +
  geom_point()
```

```{r}
c %>% 
  ggplot(aes(country,lifeExp)) +
  geom_boxplot()
```



It's difficult to see the labels so I looked up how to do the labels vertically
```{r}
c %>% 
  ggplot(aes(country,lifeExp)) +
  geom_boxplot() +
  theme(axis.text.x=element_text(angle=90))
```
Lets make it colorful!


```{r}
c %>% 
  ggplot(aes(country,lifeExp)) +
  geom_boxplot(fill="tomato") +
  theme(axis.text.x=element_text(angle=90))
```



